A new form of transcorrelated Hamiltonian inspired by range-separated DFT

### 2020-12-17 cusp_j/paper/paper.tex
conclusion almost done

Later on, Ten-No\cite{TenNo-CPL-00-a} proposed to significantly change the paradigm of Boys and Handy: instead of using a single Slater determinant and optimizing the orbitals and the Jastrow factor, one uses a more elaborate many body theory with a frozen universal Jastrow factor.  This strategy was developed using as M{\o}ller-Plesset at second order (MP2) in Refs.\onlinecite{TenNo-CPL-00-a,HinTanTen-JCP-01} and a linearised coupled cluster ansatz in later work\cite{HinTanTen-CPL-02}).  
The Jastrow factor introduced by Ten-No, developed as a linear combination of gaussian functions and referred to as the frozen gaussian geminal (FROGG), only depends on the inter-electronic coordinate and is optimized such that the scalar effective two-electron  potential cancels on average the coulomb potential near $r_{12}=0$. The use of gaussian functions to represent the Jastrow factor enabled Ten-No to develop a numerical scheme to evaluate exactly the three-electron integrals present in TC Hamiltonian\cite{TenNo-CPL-00-b}, which is an alternative to the density-fitting method proposed by Ten-No in the general context of F12 methods\cite{TenMan-JCP-03}. 
In Ref.\onlinecite{HinTanTen-JCP-01}, Ten-No \textit{et. al.} introduced a biorthogonal approach, 
which allows to treat the non-hermitian TC Hamiltonian in a mathematical form more suited 
to develop many-body perturbation theory approaches. 
It is also noteworthy that the eigenfunctions of the TC Hamiltonian are invariant by orbital rotation, 
and therefore in the limit where one gives full flexibility to the wave function, 
the biorthogonal and usual orthogonal approaches give the same results. 
Nevertheless, as approximations are necessary done in the wave function ansatz for realistic systems, 
the biorthogonal approach might improve the quality of the results.

In latter works, Umezawa \textit{et. al.}\cite{UmeTsu-JCP-03,UmeTsuOhnShiChi-JCP-05} introduced a scheme to couple TC equations with the usual variational Monte Carlo (VMC) scheme in order to optimise the Jastrow factor through a variance miminization. 
An attempt to make TC equations variational have been proposed by Luo\cite{Luo-JCP-10,Luo-JCP-11} based on the empirical experience that in the original single determinant TC approach of Boys and Handy, the orbital optimization seemed to be the source of loss of the variational property. Luo therefore proposed to replace the original TC equations to optimize the orbitals by that of a general Jastrow-Slater ansatz in VMC. 
An alternative to the non hermitian nature of TC theory have been proposed by Yanai \textit{et. al.}\cite{YanShi-JCP-12}  where the Jastrow factor is expressed as an anti-hermitian operator, allowing one to obtain an hermitian effective Hamiltonian after the similarity transformation, which is then efficiently truncated through the use the canonical transformation\cite{NeuYanCha-MolPhys-10}.

@@

The paper is organized as follows. 
In the first section, inspired by the RS-DFT effective Hamiltonian,  
we derive the analytical form of a new Jastrow factor tuned by a unique parameter $\mu$, and discuss the physical content of such a simple Jastrow factor. 
We show then that the corresponding transcorrelated Hamiltonian has a straightforward analytical structure, which is briefly reviewed as a function of the $\mu$ parameter. 
Then, we perform a numerical study of the behaviour of the ground state eigenpair of $\tilde{H}[\mu]$ in the case of the helium atom: 
in the first part we analyze the speed of convergence of the ground state energy with respect to the basis set and the parameter $\mu$, 
then we study the behaviour of the right eigenvectors in real space and compare it with the numerically exact ground state wave function. 
We show that the rapid convergence of the total energies coincides with a rapid convergence of the right eigenvectors in real space, and how to recover a very good approximation of the exact ground state wave function. 
Based on such encouraging results, we study the ground state energy of the helium isoelectronic series from Li$^+$ to Ne$^{8+}$ and compare different schemes to obtain 
a value of $\mu$ which automatically adapts to the system and still provides a fast convergence of the total energies. 
Eventually, we study the H$_2$ molecule and compare different regimes of correlation. 
Numerical comparison with the TC Hamiltonian obtained with the FROGG is also performed for all systems studied.

>>>>

Later on, Ten-No\cite{TenNo-CPL-00-a} proposed to significantly change the paradigm of Boys and Handy: instead of using a single Slater determinant and optimizing the orbitals and the Jastrow factor, one uses a more elaborate many body theory with a frozen universal Jastrow factor.  This strategy was developed using as M{\o}ller-Plesset at second order (MP2) in Refs.\onlinecite{TenNo-CPL-00-a,HinTanTen-JCP-01} and a linearised coupled cluster ansatz in later work\cite{HinTanTen-CPL-02}.  
The Jastrow factor introduced by Ten-No, developed as a linear combination of gaussian functions and referred to as the frozen gaussian geminal (FROGG), only depends on the inter-electronic coordinate and is optimized such that the scalar effective two-electron  potential cancels on average the coulomb potential near $r_{12}=0$. The use of gaussian functions to represent the Jastrow factor enabled Ten-No to develop a numerical scheme to evaluate exactly the three-electron integrals present in TC Hamiltonian\cite{TenNo-CPL-00-b}, which is an alternative to the density-fitting method proposed by Ten-No in the general context of F12 methods\cite{TenMan-JCP-03}. 
In Ref.\onlinecite{HinTanTen-JCP-01}, Ten-No \textit{et. al.} introduced a biorthogonal approach, 
which allows to treat the non-hermitian TC Hamiltonian in a mathematical form more suited 
to develop many-body perturbation theory approaches. 
It is also noteworthy that the eigenfunctions of the TC Hamiltonian are invariant by orbital rotation, 
and therefore in the limit where one gives full flexibility to the wave function, 
the biorthogonal and usual orthogonal approaches give the same results. 
Nevertheless, as approximations are necessary done in the wave function ansatz for realistic systems, 
the biorthogonal approach might improve the quality of the results in practice.

In later works, Umezawa \textit{et. al.}\cite{UmeTsu-JCP-03,UmeTsuOhnShiChi-JCP-05} introduced a scheme to couple TC equations with the usual variational Monte Carlo (VMC) scheme in order to optimise the Jastrow factor through a variance miminization. 
An attempt to make TC equations variational have been proposed by Luo\cite{Luo-JCP-10,Luo-JCP-11} based on the empirical experience that in the original single determinant TC approach of Boys and Handy, the orbital optimization seemed to be the source of loss of the variational property. Luo therefore proposed to replace the original TC equations to optimize the orbitals by that of a general Jastrow-Slater ansatz in VMC. 
An alternative to the non hermitian nature of TC theory have been proposed by Yanai \textit{et. al.}\cite{YanShi-JCP-12}  where the Jastrow factor is expressed as an anti-hermitian operator, allowing one to obtain an hermitian effective Hamiltonian after the similarity transformation, which is then efficiently truncated through the use the canonical transformation\cite{NeuYanCha-MolPhys-10}. Recent applications of such method to the context of quantum computing have been recently published\cite{ValTak-PCCP-20}.

@@

The paper is organized as follows. 
In the first section, inspired by the RS-DFT effective Hamiltonian,  
we derive the analytical form of a new Jastrow factor tuned by a unique parameter $\mu$, and discuss the physical content of such a simple Jastrow factor. 
We show then that the corresponding transcorrelated Hamiltonian has a straightforward analytical structure, which is briefly reviewed as a function of the $\mu$ parameter. 
Then, we perform a numerical study of the behaviour of the ground state eigenpair of $\tilde{H}[\mu]$ in the case of the helium atom: 
in the first part we analyze the speed of convergence of the ground state energy with respect to the basis set and the parameter $\mu$, 
then we study the behaviour of the right eigenvectors in real space and compare it with the numerically exact ground state wave function. 
We show that the rapid convergence of the total energies coincides with a rapid convergence of the right eigenvectors in real space, and how to recover a very good approximation of the exact ground state wave function. 
Based on such encouraging results, we study the ground state energy of the helium isoelectronic series from H$^-$ to Ne$^{8+}$ and compare different schemes to obtain 
a value of $\mu$ which automatically adapts to the system and still provides a fast convergence of the total energies. 
Eventually, we study the H$_2$ molecule and compare different regimes of correlation. 
Numerical comparison with the TC Hamiltonian obtained with the FROGG is also performed for all systems studied.
